Binary Knockoffs Notes

### 2015-06-20 masters_paper/notes.tex
Notes in relatively good shape

\section{Random Bernoulli Generation}
    Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values. A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$. Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $m=\diag(M)$. For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
    \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
    for all $i\neq j$\footnote{``On parametric families for sampling binary data with specified mean and correlation" - http://arxiv.org/abs/1111.0576}. Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $\bf x$.

\subsection{Gaussian Copula Family}
    Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties. There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.

\subsection{$\mu$-Conditionals family}
    It turns out that there exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as

@@

    for some monotone function $\mu:\R\to[0,1]$. This is essentially a binomial family GLM for a link function $\mu$. If one takes all of the $a_{kj}$, they can form a lower triangular matrix $A$, and then the joint density can be expressed as

@@

    If $\mu$ is chose such that it is a bijection and differentiable, there is a unique $M$ such that $\E(x_ix_i^T)=M$\footnote{``On parametric families for sampling binary data with specified mean and correlation"}. It turns out that the natural choice for $\mu$ is the logistic link function, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
    \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
    when $i\neq j$. I think this dictates that if $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are independent.  \par

@@

    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & \E(x_i) \end{array} \right]^T \]

@@

            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{\E(x_i)}{1-\E(x_i)}\right) \end{array} \right]^T \]

@@

\section{Generating Knockoffs}
    My method for generating binary knockoffs broadly involves two steps:
    \begin{enumerate}
        \item I use either the equal correlation method or the SDP method described in the original knockoff method to find $\mathbf s$ such that $\diag(\hat\Sigma)-\mathbf s$ is small and 
        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag(\mathbf s )\\ \hat\Sigma - \diag(\mathbf s) & \hat\Sigma \end{array}\right] \succeq 0 \]
        I use a subscript $L$ for large to denote items associated with the joint distribution of $[\mathbf x \; \mathbf{\tilde x}]$. If $\mathbf m_L = \E\left([\mathbf x \; \mathbf{\tilde x}]\right)^T = [\mathbf m \; \mathbf m]^T$, then the desired cross moment matrix of the joint distribution is 
        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
        To ensure that this is a valid cross moment matrix for a binary random vector is that 
        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq M_{L,ij} \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
        I've built a check into the code for this, but in practice it shouldn't be a worry. This condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$. Since the we are either keeping the value $M_{L,ij}$ from a valid cross moment matrix or minimizing $\vert M_{L,ij} - \mathbf m_{L,j}\mathbf m_{L,i}\vert$, it would be very surprising if this condition was violated.

        \item I can fit the matrix $A$ that will generate random binary variables similarly to the method described in section $3.2$ which have cross moments $X_L$. This can be used to generate the $\tilde x_i$ sequentially as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$ so as to create $\tilde X \st X$. 
    \end{enumerate}
    \subsection{More Detail on Fitting $A$}
        I could fit $A$ based on $M_L$ exactly as described in the second method of $3.2$, however, this isn't exactly what I do. First off, with $p$ being potentially large, the simulation method for estimating $f(\mathbf a)$ and $H f(\mathbf a)$ was the obvious choice. This involves fitting $\mathbf a_i$ based on the conditional distribution of $\mathbf x_{L,i}$ given randomly drawn partial vectors $\mathbf x_{L,-i}$. There is no need to simulate the marginal distribution of $\mathbf x$ though, since the simulation is only approximate and we already have a number of realizations in $X$. Thus, I only fit the lower half of $A$, using this process: 
        \begin{enumerate}
            \item To get a simulation of size at least $K$, I create a matrix $X_F$ ($F$ for fixed) which is initially $X$ stacked up until it has $K'\geq K$ rows. 
            \item For each $p<i\leq 2p$,
                \begin{itemize}
                    \item Where $\mathbf x_F^{(k)}$ is the $k$th row of $X_F$, the rows $\mathbf a_i$ are fit sequentially by Newton-Raphson iterations with 
                    \[f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right] \]
                    and
                    \[H f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
                    \item If the iterations won't converge, I attempt to $f(\mathbf a_i)=m_i^*(\tau)$ for increasing values of $\tau$ until it converges.
                    \item After $\mathbf a_i$ is fit, a new column $X_i$ is drawn as independent Bernoulli with probability vector $\mu(X_F \mathbf a_i)$.   
                    \item $X_F$ is updated to 
                    \[ X_F = [X_F X_i] \]
                \end{itemize}
            \item At the end, the first $n$ rows of $X_F$ are taken as $[X \tilde X]$, though the rows corresponding to any copy of $X$ would work equally well.
        \end{enumerate}
        Some thoughts/concerns/explanations:
        \begin{itemize}
            \item Since $x_{-i}^{(k)}$ is replaced with $x_F^{(k)}$, which is a hybrid of real and simulated data, I am not sure there is the same theoretical guarantee that a unique $A$ matrix exists. In practice though it still worked pretty well, and has the advantage that we are deriving $\tilde X$ such that $\E(X'\tilde X) = \hat \Sigma- \diag(\mathbf s)$.
            \item A hybrid of simulating $x_{-i}^{(k)}$ and keeping fixed $x_F^{(k)}$ would be to draw from the rows of $X$ $K$ times, with replacement, then simulating the rest of the $x_{-i}^{(k)}$ vector. I'm not convinced there is a good reason to do this. 
            \item By not simulating, there is the advantage of not needing to fit the upper half of $A$.
            \item By not redrawing $x_{-i}^{(k)}$ each time, there is less computation for the computer. As well, the multiplication
                \[ \left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
            needn't be redone for each iteration, and only partially recalculated for each $i$ (though I don't have this implemented yet). The downside might be that error is getting compounded over each $i$.
        \end{itemize}

@@

\subsection{Comparison to Original Knockoffs in LASSO}
The basic idea is to compare how the two sorts of knockoffs compare when $X$ is binary and we are fitting a LASSO regression. Do they select the same variables? Do the binary knockoffs control FDR more or less conservatively? Situations to model:
\begin{itemize}
    \item $X$ arises out of the assumed Isling model, with no high order interactions. 
    \item $X$ does not arise out of this model, and does have higher order interactions. Could possibly model this by drawing the vector $\mathbf x$ or subsets of it as multinomial with probabilities from a Dirichlet distribution.
    \item $X$ is from a real world data set, likely something in genetics.
    \item $y$ is normally distributed around $X\beta$.
    \item $y$ is drawn from some other distribution, possibly skewed, heavy tailed, or light tailed.
    \item $y$ is from a real world data set.
\end{itemize}

\subsection{Evaluation of Binary Knockoffs in other GLMS}
The one of the most interest would be logits. It would be good to see how successful the binary knockoffs are in controlling FDR without the theoretical guarantees that LASSO provides.
\begin{itemize}
    \item $X$ arises out of the Isling model, in which case it seems like the knockoffs should control FDR.
    \item $X$ has higher order interactions, which might make the knockoffs perform poorly.
    \item $X$ is from a real world data set, likely something in genetics.
    \item $y$ is simulated based on the assumptions of the model.
    \item $y$ is simulated to violate assumptions of the model.
    \item $y$ is from a real world data set.
\end{itemize}

\subsection{Generalizations/Harebrained Ideas}
I have two ideas for extensions if I have enough time.
\begin{enumerate}
    \item Still with binary data, one might be able to simulate binary variables with higher order interactions in the generation of $X$ by including higher order interactions in the regression of $X_i$ on $X_{-i}$. This would lead to $A$ being a matrix of higher dimension. Even if this worked, extending the method to knockoffs may not be obvious.
    \item In the general case where $X$ is not binary, I can almost imagine a method set up along similar lines to the binary case.
    \begin{itemize}
        \item The desired covariance matrix would arise as before.
        \item Each $x_i$'s marginal would be approximated by a kernel density estimate on $X_i$.
        \item A $x_i \st x_1, \ldots, x_{i-1}$ would be drawn from some reweighing of this marginal to achieve the proper covariance. For instance, if $F_i^{-1}$ is the inverse CDF of the marginal kernel density for $x_i$ and $u_i\st x_1, \ldots, x_{i-1}$ is a RV on $(0,1)$, then $x_i=F_i^{-1}(u_i)$.
    \end{itemize}
\end{enumerate}

>>>>

\section{Random Bernoulli Generation}
    Thankfully, there has been a reasonable amount of work on how one can generate random Bernoulli vectors with some kind of correlation among among the values. A random Bernoulli vector $\bf x$ can be summarized by its first two moments: a mean vector $\E(\bf x)=m\in(0,1)^p$ and cross-moment matrix $\E(\bf x\bf x^T)=M\in(0,1)^{p\times p}$. Obviously, $m_i = \p(x_i=1)$, $M_{ij} = \p(x_i=x_j=1)$, and $\mathbf m=\diag(M)$. For an arbitrary symmetric $M$ to be valid cross-moment matrix, $M-mm^T\succeq0$, and
    \[ \max\{0,m_i+m_j -1\} \leq M_{ij} \leq \min\{m_i,m_j\}\]
    for all $i\neq j$\footnote{``On parametric families for sampling binary data with specified mean and correlation" - http://arxiv.org/abs/1111.0576}. Given a qualifying $M$, or observed $X$, there are a few ways of generating more random $\bf x$.

\subsection{Gaussian Copula Family}
    Since multivariate normal distributions are easy to randomly draw, the idea is to find some random normal variable $\mathbf z\sim N_p({\bf 0},\Sigma)$ such that,for $x_i = I(z_i<0)$, $x$ has the desired properties. There are a number of ways to do this\footnote{``On the Generation of Correlated Artificial Binary Data" - http://epub.wu.ac.at/286/1/document.pdf}\footnote{``On parametric families for sampling binary data with specified mean and correlation"}, but it turns out that there is only certain to exist a working $\Sigma$ in the bivariate case.

\subsection{$\mu$-Conditionals family}
    There exists a more flexible family which will always work for arbitrary $M$ called $\mu$-conditionals. The basic idea is that the $X$ is generate sequentially as

@@

    for some monotone function $\mu:\R\to(0,1)$. This is essentially a binomial family GLM for a link function $\mu$. If one takes all of the $a_{kj}$, they can form a lower triangular matrix $A$, and then the joint density can be expressed as

@@

    If $\mu$ is chosen such that it is a bijection and differentiable, there is a unique $A$ such that $\E(\mathbf x \mathbf x^T)=M$ when generated from this model\footnote{``On parametric families for sampling binary data with specified mean and correlation"}. The natural choice for $\mu$ is the logistic link function, which yields the Ising model, the ``binary analogue of the multivariate normal distribution which is the maximum entropy distribution on $\R^p$ having a given covariance matrix." Additionally, it has the usual benefit that the coefficients can be viewed as a log odds ratio:
    \[a_{ij} = \log\left(\frac{\p(x_j=x_k=1)\p(x_j=x_k=0)}{\p(x_j=0,x_k=1)\p(x_j=1,x_k=0)}\right) \]
    when $i\neq j$. I think this dictates that if $\bf x$ is generated from this model with $a_{ij}=0$, then $x_i$ and $x_j$ are independent.  \par

@@

    \[\mathbf m_i^*(\tau) = (1-\tau)\mathbf m_i + \tau \left[ \begin{array}{cccc} 0 & \ldots & 0 & M_{ii} \end{array} \right]^T \]

@@

            \[\mathbf a_i = \left[ \begin{array}{cccc} 0 & \ldots & 0 & \log\left(\frac{M_{ii}}{1-M_{ii}}\right) \end{array} \right]^T \]

@@

\section{Generating Knockoffs}
    My method for generating binary knockoffs broadly involves two steps:
    \begin{enumerate}
        \item I use either the equal correlation method or the SDP method described in the original knockoff paper to find $\mathbf s$ such that $\|\diag(\hat\Sigma)-\mathbf s\|$ is small and 
        \[ \Sigma_L = \left[\begin{array}{cc}  \hat\Sigma & \hat\Sigma - \diag(\mathbf s )\\ \hat\Sigma - \diag(\mathbf s) & \hat\Sigma \end{array}\right] \succeq 0 \]
        I use a subscript $L$ for large to denote items associated with the joint distribution of $[\mathbf x \; \mathbf{\tilde x}]$. If $\mathbf m_L = \E\left([\mathbf x \; \mathbf{\tilde x}]\right)^T = [\mathbf m \; \mathbf m]^T$, then the desired cross moment matrix of the joint distribution is 
        \[ M_L = \Sigma_L + \mathbf m_L\mathbf m_L^T\]
        To ensure that this is a valid cross moment matrix for a binary random vector is that 
        \[\max\{0,\mathbf m_{L,i} + \mathbf m_{L,j} -1\} \leq M_{L,ij} \leq \min\{\mathbf m_{L,i}, \mathbf m_{L,j}\} \]
        I've built a check into the code for this, but in practice it shouldn't be a worry. This condition is always satisfied in a neighborhood of $M_{L,ij} = \mathbf m_{L,j}\mathbf m_{L,i}$. Since the we are either keeping the value $M_{L,ij}$ from a valid cross moment matrix or minimizing $\vert M_{L,ij} - \mathbf m_{L,j}\mathbf m_{L,i}\vert$, it would be very surprising if this condition was violated.

        \item I can fit the matrix $A$ that will generate random binary variables similarly to the method described in section $3.2$ which have cross moments $X_L$. This can be used to generate the $\tilde x_i$ sequentially as $\tilde x_i \st \mathbf x,\tilde x_1,\ldots,\tilde x_{i-1}$ so as to create $\tilde X \st X$. 
    \subsection{More Detail on Fitting $A$}
        I could fit $A$ based on $M_L$ exactly as described in the second method of $3.2$, however, this isn't exactly what I do. First off, with $p$ being potentially large, the simulation method for estimating $f(\mathbf a)$ and $H f(\mathbf a)$ was the obvious choice. This involves fitting $\mathbf a_i$ based on the conditional distribution of $\mathbf x_{L,i}$ given randomly drawn partial vectors $\mathbf x_{L,-i}$. There is no need to simulate the marginal distribution of $\mathbf x$ though, since the simulation is only approximate and we already have a number of realizations in $X$. Thus, I only fit the lower half of $A$, using this process: 
            \item To get a simulation of size at least $K$, I create a matrix $X_F$ ($F$ for fixed) which is initially $X$ stacked up until it has $K'\geq K$ rows. 
            \item For each $p<i\leq 2p$,
                \begin{itemize}
                    \item Where $\mathbf x_F^{(k)}$ is the $k$th row of $X_F$, the rows $\mathbf a_i$ are fit sequentially by Newton-Raphson iterations with 
                    \[f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right] \]
                    \[H f\left(\mathbf a_i\right) \approx \frac{1}{K'}\sum_{k=1}^{K'} \mu'\left(\mathbf{a}_i^T\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\right)\left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
                    \item If the iterations won't converge, I attempt to fit $f(\mathbf a_i)=m_i^*(\tau)$ instead for increasing values of $\tau$ until it converges.
                    \item After $\mathbf a_i$ is fit, a new column $X_i$ is drawn as independent Bernoulli with probability vector $\mu(X_F \mathbf a_i)$.   
                    \item $X_F$ is updated to 
                    \[ X_F = [X_F \; X_i] \]
                \end{itemize}
            \item At the end, the first $n$ rows of $X_F$ are taken as $[X \; \tilde X]$, though the rows corresponding to any copy of $X$ would work equally well, since these all should have the desired distribution.
        Some thoughts/concerns/explanations:
        \begin{itemize}
            \item Since $x_{-i}^{(k)}$ is replaced with $x_F^{(k)}$, which is a hybrid of real and simulated data, I am not sure there is the same theoretical guarantee that a unique $A$ matrix exists. In practice though it still worked pretty well, and has the advantage that we are deriving $\tilde X$ such that $\E(X'\tilde X) = \hat \Sigma- \diag(\mathbf s)$.
            \item A hybrid of simulating $x_{-i}^{(k)}$ and keeping fixed $x_F^{(k)}$ would be to draw from the rows of $X$ $K$ times, with replacement, then simulating the rest of the $x_{-i}^{(k)}$ vector. I'm not convinced there is a good reason to do this. 
            \item By not simulating, there is the advantage of not needing to fit the upper half of $A$.
            \item By not redrawing $x_{-i}^{(k)}$ each time, there is less computation for the computer. As well, the multiplication
                \[ \left[\begin{array}{c} \mathbf x_F^{(k)} \\ 1 \end{array}\right]\left[\begin{array}{cc} \mathbf [x_F^{(k)}]^T & 1 \end{array}\right] \]
            needn't be redone for each iteration, and only partially recalculated for each $i$ (though I don't have this implemented yet). The downside might be that error is getting compounded over each $i$.
        \end{itemize}

@@

\subsection{Comparison to Original Knockoffs in LASSO}
The basic idea is to compare how the two sorts of knockoffs compare when $X$ is binary and we are fitting a LASSO regression. Do they select the same variables? Do the binary knockoffs control FDR more or less conservatively? Situations to test:
\begin{itemize}
    \item $X$ arises out of the assumed Isling model, with no high order interactions. 
    \item $X$ does not arise out of this model, and does have higher order interactions. Could possibly model this by drawing the vector $\mathbf x$ or subsets of it as multinomial with probabilities from a Dirichlet distribution.
    \item $X$ is from a real world data set, likely something in genetics.
    \item $y$ is normally distributed around $X\beta$.
    \item $y$ is drawn from some other distribution, possibly skewed, heavy tailed, or light tailed.
    \item $y$ is from a real world data set.
\end{itemize}

\subsection{Evaluation of Binary Knockoffs in other GLMS}
The one of the most interest would be logits. It would be good to see how successful the binary knockoffs are in controlling FDR without the theoretical guarantees that LASSO provides.
\begin{itemize}
    \item $X$ arises out of the Isling model, in which case it seems like the knockoffs should control FDR.
    \item $X$ has higher order interactions, which might make the knockoffs perform poorly.
    \item $X$ is from a real world data set, likely something in genetics.
    \item $y$ is simulated based on the assumptions of the regression model.
    \item $y$ is simulated to violate assumptions of the regression model.
    \item $y$ is from a real world data set.
\end{itemize}

\subsection{Generalizations/Harebrained Ideas}
I have two seeds of ideas for extensions if I have enough time.
\begin{enumerate}
    \item Still with binary data, one might be able to simulate binary variables with higher order interactions in the generation of $X$ by including higher order interactions in the regression of $X_i$ on $X_{-i}$. This would lead to $A$ being a matrix of higher dimension. Even if this worked, extending the method to knockoffs may not be obvious.
    \item In the general case where $X$ is not binary, I can almost imagine a method set up along similar lines to the binary case.
    \begin{itemize}
        \item The desired covariance matrix could be chosen as in the original knockoff paper.
        \item Each $x_i$'s marginal distribution would be approximated by a kernel density estimate on $X_i$.
        \item A $x_i \st x_1, \ldots, x_{i-1}$ would be drawn from some reweighing of this marginal to achieve the proper covariance. For instance, if $F_i^{-1}$ is the inverse CDF of the marginal kernel density for $x_i$ and $u_i\st x_1, \ldots, x_{i-1}$ is a RV on $(0,1)$, then $x_i=F_i^{-1}(u_i)$.
        \item Maybe fit generalized additive model for each successive $x_i$ with kernel regression for $x_1,\ldots,x_{i-1}$ to predict mean. Then, skew marginal of $x_i$ until it has that mean. 
    \end{itemize}
\end{enumerate}
